---
  title: "NFL Score Prediction Model: Feature Engineering"
author: "Wade Fuller"
date: "03/03/2021"
output:
  pdf_document: default
html_document: default
---

```{r setup, include=FALSE}
rm(list = ls())
library(knitr)
knitr::opts_chunk$set(cache = TRUE, echo = TRUE,
                      message = FALSE, dpi = 180,
                      fig.width = 8, fig.height = 5)
knitr::opts_knit$set(root.dir = "~/Documents/Scripts/R/pred/")
wd <- "/Users/wade/Documents/Scripts/R"
library(tidymodels)
library(tidyverse)
library(nflfastR)
library(ggthemes)
library(zoo)
library(doParallel)
library(vip)
theme_set(theme_minimal(base_size = 10))
```

## Introduction
Trying to make some general data tables for each team with the performance metrics and model features that I care about for each week of a given season.

```{r init}
games <- readRDS(url("http://www.habitatring.com/games.rds"))
games %>% arrange(-season,gameday)
```
This data frame contains every game since 1999 as well as a treasure trove of metadata about each game.

### Split data
The data is imported such that each record represents one game. This means that we have information about both the home and away team on the same row To make this data easier to manipulate, we have to split this data into home and away frames.

```{r split_data}
home_games <- games %>%
  filter(season >= 2000) %>%
  transmute(
    game_id
    ,gameday
    ,team = home_team
    ,opposing_team = away_team
    ,points_for = home_score
    ,points_against = away_score
    ,result
    ,total_points_scored = total
    ,pct_points_scored = points_for/total_points_scored
    ,season
    ,week
    ,gameday
    ,weekday
    ,gametime
    ,team_rest = home_rest
    ,opponent_rest = away_rest
    ,spread_line = spread_line * -1
    ,total_line
    ,home_spread_odds
    ,away_spread_odds
    ,home_moneyline
    ,away_moneyline
    ,impl_prob_home = if_else(sign(home_moneyline)== -1,(-1*home_moneyline/((-1*home_moneyline)+100)),(100/(home_moneyline+100)))
    ,impl_prob_away = if_else(sign(away_moneyline)== -1,(-1*away_moneyline/((-1*away_moneyline)+100)),(100/(away_moneyline+100)))
    ,div_game
    ,home = 1
  )
away_games <- games %>%
  filter(season >= 2000) %>%
  transmute(
    game_id
    ,gameday
    ,team = away_team
    ,opposing_team = home_team
    ,points_for = away_score
    ,points_against = home_score
    ,result = result * -1
    ,total_points_scored = total
    ,pct_points_scored = points_for/total_points_scored
    ,season
    ,week
    ,gameday
    ,weekday
    ,gametime
    ,team_rest = away_rest
    ,opponent_rest = home_rest
    ,spread_line
    ,total_line
    ,home_spread_odds
    ,away_spread_odds
    ,home_moneyline
    ,away_moneyline
    ,impl_prob_home = if_else(sign(home_moneyline)== -1,(-1*home_moneyline/((-1*home_moneyline)+100)),(100/(home_moneyline+100)))
    ,impl_prob_away = if_else(sign(away_moneyline)== -1,(-1*away_moneyline/((-1*away_moneyline)+100)),(100/(away_moneyline+100)))
    ,div_game
    ,home = 0
  )
game_df <- bind_rows(home_games, away_games) %>%
  mutate_if(is.character, factor) %>%
  mutate(home = as.factor(home))
game_df <- game_df %>% mutate(
  win = if_else(game_df$result > 0,1,0) # fully recognizing that this misattributes ties to losses.
  ,spread_adjusted_result = spread_line + result
  ,spread_cover = if_else(spread_line + result > 0,1,0)
  ,impl_win_prob = if_else(home == 1,impl_prob_home,impl_prob_away)
  ,game_info_gain = if_else(win == 1,pmax(1-2*impl_win_prob^3,0, na.rm = TRUE),pmax(1-2*impl_win_prob^3-1,-1, na.rm = TRUE))
  ,win_prob_captured = if_else(win == 1, impl_win_prob,-impl_win_prob)
  ,rest_diff = team_rest - opponent_rest
)
game_df %>% filter(season == 2021, team == 'DAL') %>% arrange(week)
```

### Generate features

```{r feature_set}
trending_fs <- game_df %>%
  arrange(season,team,week) %>%
  group_by(season,team) %>%
  mutate(win_pct_avg = rollapply(win, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
         ,point_diff_avg = rollapply(result, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
         ,points_for_avg = rollapply(points_for, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
         ,win_pct_ats_avg = rollapply(spread_cover, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
         ,point_diff_ats_avg = rollapply(spread_adjusted_result, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
         ,impl_win_prob_avg = rollapply(impl_win_prob, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
         ,game_info_gain_avg = rollapply(game_info_gain, list(-(16:1)), sum, align = 'right' , fill = NA, partial = T)
         ,win_prob_captured_avg = rollapply(win_prob_captured, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
         ,win_pct_trend = rollapply(win, list(-(4:1)), mean, align = 'right' , fill = NA, partial = T)
         ,point_diff_trend = rollapply(result, list(-(4:1)), mean, align = 'right' , fill = NA, partial = T)
         ,points_for_trend = rollapply(points_for, list(-(4:1)), mean, align = 'right' , fill = NA, partial = T)
         ,win_pct_ats_trend = rollapply(spread_cover, list(-(4:1)), mean, align = 'right' , fill = NA, partial = T)
         ,point_diff_ats_trend = rollapply(spread_adjusted_result, list(-(4:1)), mean, align = 'right' , fill = NA, partial = T)
         ,impl_win_prob_trend = rollapply(impl_win_prob, list(-(4:1)), mean, align = 'right' , fill = NA, partial = T)
         ,game_info_gain_trend = rollapply(game_info_gain, list(-(4:1)), sum, align = 'right' , fill = NA, partial = T)
         ,win_prob_captured_trend = rollapply(win_prob_captured, list(-(4:1)), mean, align = 'right' , fill = NA, partial = T)) %>%
  select(game_id
         ,season
         ,week
         ,team
         ,win_pct_avg
         ,point_diff_avg
         ,points_for_avg
         ,win_pct_ats_avg
         ,point_diff_ats_avg
         ,impl_win_prob_avg
         ,game_info_gain_avg
         ,win_prob_captured_avg
         ,win_pct_trend
         ,point_diff_trend
         ,points_for_trend
         ,win_pct_ats_trend
         ,point_diff_ats_trend
         ,impl_win_prob_trend
         ,game_info_gain_trend
         ,win_prob_captured_trend
  )

pythag_df <- game_df %>%
  group_by(season,team) %>%
  arrange(season,team,week) %>%
  mutate(gms = row_number()-1
         ,agg_pts_for = cumsum(points_for)
         ,agg_pts_against = cumsum(points_against)
         ,pts_for_entering = lag(agg_pts_for)
         ,pts_against_entering = lag(agg_pts_against)) %>%
  mutate(pythag_win_pct = (pts_for_entering^2.37)/((pts_for_entering^2.37)+(pts_against_entering^2.37))
         ,pythag_wins = pythag_win_pct*gms) %>%
  ungroup() %>%
  select(game_id,team,pythag_win_pct,pythag_wins,gms)

fs_blend <- trending_fs %>% inner_join(pythag_df,by=c('game_id'='game_id','team'='team'))
```

``` {r}
fs_blend %>% 
  select(-matches("_trend")) %>% 
  filter(season == 2021, week == 18) %>% 
  arrange(week,desc(pythag_win_pct))
```

``` {r}
## INITIAL PBP LOAD
# pbp <- nflreadr::load_pbp(seasons = 2000:2021)

# plys <- readRDS(paste0(wd,"nflfastscrapr_plys.RData"))
pbp <- readRDS("~/Documents/Scripts/R/pred/data/pbp_2000_2021.rds")

# View latest games with pbp data
gms_w_pbp <- pbp %>% distinct(game_id) %>% arrange(desc(game_id))

# View latest games with results
gms_w_result <- games %>% filter(!is.na(result)) %>% distinct(game_id) %>% arrange(desc(game_id))

# Define games to fetch via anti_join
gms_to_fetch <- gms_w_result %>%
  filter(substr(game_id,0,4) >= 2019) %>%
  anti_join(gms_w_pbp)

# Load the new pbp data
new_pbp <- nflfastR::build_nflfastR_pbp(gms_to_fetch)

# Check the new games
new_pbp %>% distinct(game_id)

# Combine new rows and existing ones
pbp_adj <- rbind(pbp,new_pbp)
pbp_adj %>% distinct(game_id) %>% arrange(desc(game_id))
pbp <- as.data.frame(pbp_adj)
saveRDS(pbp, file = "~/Documents/Scripts/R/pred/data/pbp_2000_2021.RDS")
```

``` {r}
plys <- pbp %>%
  dplyr::mutate(
    home_team = gsub("JAC", "JAX", home_team),
    away_team = gsub("JAC", "JAX", away_team),
    home_team = gsub("SD", "LAC", home_team),
    away_team = gsub("SD", "LAC", away_team),
    home_team = gsub("OAK", "LV", home_team),
    away_team = gsub("OAK", "LV", away_team),
    home_team = gsub("STL", "LA", home_team),
    away_team = gsub("STL", "LA", away_team)
  )

plys %>%
  filter(game_id == '2021_01_DAL_TB',!is.na(epa), !is.na(ep), !is.na(posteam), !is.na(down), play_type == "pass" | play_type == "run" | penalty == 1, qb_kneel != 1) %>%
  group_by(defteam) %>%
  summarise( n()
             ,sum(qb_epa)
             ,mean(qb_epa)
             ,mean(epa)
             ,sum(epa)
             ,sum(ep)
             ,mean(ep))
plys %>%
  filter(game_id == '2021_01_DAL_TB',!is.na(epa), !is.na(ep), !is.na(posteam), !is.na(down), play_type == "pass" | play_type == "run" | penalty == 1, qb_kneel != 1) %>%
  group_by(game_id,season,week,posteam) %>%
  summarise(
    off_epa = mean(epa)
    ,off_pass_epa = mean(epa[qb_dropback == 1])
    ,off_rush_epa = mean(epa[qb_dropback == 0])
    ,success_rate = mean(success[!is.na(success)])
    ,cpoe=mean(cpoe[!is.na(cpoe)])
    ,qb_epa=mean(qb_epa[!is.na(qb_epa)])
    ,.groups = 'drop'
  ) %>%
  arrange(posteam,week) %>%
  rename(team = posteam)


init_oepa_df <- plys %>%
  filter(!is.na(epa), !is.na(ep), !is.na(posteam), !is.na(down)
         ,play_type == "pass" | play_type == "run" | penalty == 1, qb_kneel != 1) %>%
  group_by(game_id,season,week,posteam) %>%
  summarise(
    off_epa = mean(epa)
    ,off_pass_epa = mean(epa[qb_dropback == 1])
    ,off_rush_epa = mean(epa[qb_dropback == 0])
    ,success_rate = mean(success[!is.na(success)])
    ,cpoe=mean(cpoe[!is.na(cpoe)])
    ,qb_epa=mean(qb_epa[!is.na(qb_epa)])
    ,.groups = 'drop'
  ) %>%
  arrange(posteam,week) %>%
  rename(team = posteam)

middle_epa_df <- fs_blend %>% select(-c(win_pct_avg:pythag_wins)) %>%
  left_join(init_oepa_df)

f_epa_df <- middle_epa_df %>%
  group_by(team) %>%
  arrange(team,season,week) %>%
  mutate(
    ,r1_epa_pp = rollapply(off_epa, list(-(1:1)), mean, align = 'right' , fill = NA, partial = T), .after = team
    ,r2_epa_pp = rollapply(off_epa, list(-(2:1)), mean, align = 'right' , fill = NA, partial = T)
    ,r4_epa_pp = rollapply(off_epa, list(-(4:1)), mean, align = 'right' , fill = NA, partial = T)
    ,r4_pass_epa = rollapply(off_pass_epa, list(-(4:1)), mean, align = 'right' , fill = NA, partial = T)
    ,r4_rush_epa = rollapply(off_rush_epa, list(-(4:1)), mean, align = 'right' , fill = NA, partial = T)
    ,r4_qb_epa = rollapply(qb_epa, list(-(4:1)), mean, align = 'right' , fill = NA, partial = T)
    ,r4_success_rt = rollapply(success_rate, list(-(4:1)), mean, align = 'right' , fill = NA, partial = T)
    ,r12_epa_pp = rollapply(off_epa, list(-(12:1)), mean, align = 'right' , fill = NA, partial = T)
    ,r12_pass_epa = rollapply(off_pass_epa, list(-(12:1)), mean, align = 'right' , fill = NA, partial = T)
    ,r12_rush_epa = rollapply(off_rush_epa, list(-(12:1)), mean, align = 'right' , fill = NA, partial = T)
    ,r12_qb_epa = rollapply(qb_epa, list(-(12:1)), mean, align = 'right' , fill = NA, partial = T)
    ,r12_success_rt = rollapply(success_rate, list(-(12:1)), mean, align = 'right' , fill = NA, partial = T)
    ,r16_epa_pp = rollapply(off_epa, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
    ,r16_pass_epa = rollapply(off_pass_epa, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
    ,r16_rush_epa = rollapply(off_rush_epa, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
    ,r16_qb_epa = rollapply(qb_epa, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
    ,r16_success_rt = rollapply(success_rate, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
  ) %>%
  ungroup() %>%
  select(-game_id)

sea_epa_df <- middle_epa_df %>%
  group_by(team,season) %>%
  arrange(team,season,week) %>%
  mutate(
     sea_epa_pp = rollapply(off_epa, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
    ,sea_pass_epa = rollapply(off_pass_epa, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
    ,sea_rush_epa = rollapply(off_rush_epa, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
    ,sea_qb_epa = rollapply(qb_epa, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
    ,sea_success_rt = rollapply(success_rate, list(-(16:1)), mean, align = 'right' , fill = NA, partial = T)
  ) %>%
  ungroup() %>%
  select(-game_id)

# Create a final EPA DF
sea_epa_ntiles <- sea_epa_df %>% 
  filter(season >= 2010) %>% 
  arrange(-sea_epa_pp) %>% 
  mutate(
    epa_pp_ntile = ntile(sea_epa_pp,100)
   ,pass_epa_ntile = ntile(sea_pass_epa,100)
   ,rush_epa_ntile = ntile(sea_rush_epa,100)
   ,qb_epa_ntile = ntile(sea_qb_epa,100)
   ,success_rt_ntile = ntile(sea_success_rt,100)
  )

fs_final <- fs_blend %>%
  left_join(sea_epa_ntiles,by = c('season','team','week')) %>%
  left_join(f_epa_df,by = c('season','team','week'))
```

``` {r}
fs_final %>% select(-(game_id)) %>% filter(week == 1 , season == 2021) %>% arrange(-epa_pp_ntile)
fs_final %>% str()
```


### Blend Features w/ Original Data Frame

```{r}
# game_features_df <- game_df %>%
#   inner_join(fs_blend,by=c('game_id'='game_id','team'='team','season'='season','week'='week'))
game_features_df <- game_df %>%
  inner_join(fs_final,by=c('game_id'='game_id','team'='team'), keep = FALSE) %>%
  rename(season = season.x, week = week.x) %>%
  select(-c(season.y,week.y))
home_df <- game_features_df %>%
  filter(home == 1)
away_df <- game_features_df %>%
  filter(home == 0)
game_w_features <- home_df %>% 
  inner_join(away_df,by=c('game_id'='game_id')) %>%
  transmute(
    # BASICS
    game_id
    ,gameday = gameday.x
    ,season = season.x
    ,week = week.x
    ,home_team = team.x
    ,away_team = opposing_team.x 
    ,spread_line = spread_line.x
    
    # RESULT
    ,home_points = points_for.x 
    ,away_points = points_against.x
    ,total_points_scored = total_points_scored.x 
    ,home_result = result.x 
    ,away_result = result.y
    ,home_win = win.x
    ,total_line = total_line.x
    
    # RESULT DETAILS
    ,spread_adj_result = spread_adjusted_result.x
    ,home_spread_cover = spread_cover.x
    ,away_spread_cover = spread_cover.y
    
    # TEAM PERFORMANCE METADATA
    ## WIN/LOSS RECORD, POINT DIFF VS EXPECTATION
    ,home_win_pct_season = win_pct_avg.x
    ,away_win_pct_season = win_pct_avg.y
    ,win_pct_season_diff = win_pct_avg.x - win_pct_avg.y
    ,home_point_diff_season = point_diff_avg.x
    ,away_point_diff_season = point_diff_avg.y
    ,point_diff_season_diff = point_diff_avg.x - point_diff_avg.y
    ,home_win_pct_ats_season = win_pct_ats_avg.x
    ,away_win_pct_ats_season = win_pct_ats_avg.y
    ,win_pct_ats_season_diff = win_pct_ats_avg.x - win_pct_ats_avg.y
    ,home_point_diff_ats_season = point_diff_ats_avg.x
    ,away_point_diff_ats_season = point_diff_ats_avg.y
    ,point_diff_ats_season_diff = point_diff_ats_avg.x - point_diff_ats_avg.y
    
    ## WIN/LOSS TREND
    ,home_win_pct_trend = win_pct_trend.x
    ,away_win_pct_trend = win_pct_trend.y
    ,win_pct_trend_diff = win_pct_trend.x - win_pct_trend.y
    ,home_point_diff_trend = point_diff_trend.x
    ,away_point_diff_trend = point_diff_trend.y
    ,point_diff_trend_diff = point_diff_trend.x - point_diff_trend.y
    
    ## MARKET IMPLIED WIN PCT
    ,home_impl_win_prob_avg = impl_win_prob_avg.x
    ,away_impl_win_prob_avg = impl_win_prob_avg.y
    ,home_game_info_gain_avg = game_info_gain_avg.x
    ,away_game_info_gain_avg = game_info_gain_avg.y
    ,home_win_prob_captured_avg = win_prob_captured_avg.x
    ,away_win_prob_captured_avg = win_prob_captured_avg.y
    ,home_impl_win_prob_trend = impl_win_prob_trend.x
    ,away_impl_win_prob_trend = impl_win_prob_trend.y
    ,impl_win_prob_trend_diff = impl_win_prob_trend.x - impl_win_prob_trend.y
    ,home_game_info_gain_trend = game_info_gain_trend.x
    ,away_game_info_gain_trend = game_info_gain_trend.y
    ,game_info_gain_trend_diff = game_info_gain_trend.x - game_info_gain_trend.y
    ,home_win_prob_captured_trend = win_prob_captured_trend.x
    ,away_win_prob_captured_trend = win_prob_captured_trend.y
    ,win_prob_captured_trend_diff = win_prob_captured_trend.x - win_prob_captured_trend.y
    
    ## PYTHAG EXPECTATION
    ,home_pythag_win_pct = pythag_win_pct.x
    ,away_pythag_win_pct = pythag_win_pct.y
    ,pythag_win_pct_diff = pythag_win_pct.x - pythag_win_pct.y
    ,home_pythag_wins = pythag_wins.x
    ,away_pythag_wins = pythag_wins.y
    ,pythag_win_diff = pythag_wins.x - pythag_wins.y
    
    ## SPORTSBOOK ODDS
    ,home_moneyline = home_moneyline.x
    ,away_moneyline = away_moneyline.x
    ,moneyline_diff = home_moneyline.x - away_moneyline.x
    ,home_spread_odds = home_spread_odds.x
    ,away_spread_odds = away_spread_odds.x
    ,spread_odds_diff = home_spread_odds.x - away_spread_odds.x
    ,home_impl_win_prob = impl_win_prob.x
    ,away_impl_win_prob = impl_win_prob.y
    ,impl_win_prob_diff = impl_win_prob.x - impl_win_prob.y
    
    ## PLAY-BY-PLAY ANALYSIS
    ,home_r1_epa_pp = r1_epa_pp.x
    ,home_r2_epa_pp = r2_epa_pp.x
    ,home_r4_epa_pp = r4_epa_pp.x
    ,home_r4_pass_epa = r4_pass_epa.x
    ,home_r4_rush_epa = r4_rush_epa.x
    ,home_r4_qb_epa = r4_qb_epa.x
    ,home_r4_success_rt = r4_success_rt.x
    ,home_r12_epa_pp = r12_epa_pp.x
    ,home_r12_pass_epa = r12_pass_epa.x
    ,home_r12_rush_epa = r12_rush_epa.x
    ,home_r12_qb_epa = r12_qb_epa.x
    ,home_r12_success_rt = r12_success_rt.x
    ,home_r16_epa_pp = r16_epa_pp.x
    ,home_r16_pass_epa = r16_pass_epa.x
    ,home_r16_rush_epa = r16_rush_epa.x
    ,home_r16_qb_epa = r16_qb_epa.x
    ,home_r16_success_rt = r16_success_rt.x
    ,home_epa_pp_ntile= epa_pp_ntile.x
    ,home_pass_epa_ntile = pass_epa_ntile.x
    ,home_rush_epa_ntile = rush_epa_ntile.x
    ,home_qb_epa_ntile = qb_epa_ntile.x
    ,home_success_rt_ntile = success_rt_ntile.x
    
    ,away_r1_epa_pp = r1_epa_pp.y
    ,away_r2_epa_pp = r2_epa_pp.y
    ,away_r4_epa_pp = r4_epa_pp.y
    ,away_r4_pass_epa = r4_pass_epa.y
    ,away_r4_rush_epa = r4_rush_epa.y
    ,away_r4_qb_epa = r4_qb_epa.y
    ,away_r4_success_rt = r4_success_rt.y
    ,away_r12_epa_pp = r12_epa_pp.y
    ,away_r12_pass_epa = r12_pass_epa.y
    ,away_r12_rush_epa = r12_rush_epa.y
    ,away_r12_qb_epa = r12_qb_epa.y
    ,away_r12_success_rt = r12_success_rt.y
    ,away_r16_epa_pp = r16_epa_pp.y
    ,away_r16_pass_epa = r16_pass_epa.y
    ,away_r16_rush_epa = r16_rush_epa.y
    ,away_r16_qb_epa = r16_qb_epa.y
    ,away_r16_success_rt = r16_success_rt.y
    ,away_epa_pp_ntile= epa_pp_ntile.y
    ,away_pass_epa_ntile = pass_epa_ntile.y
    ,away_rush_epa_ntile = rush_epa_ntile.y
    ,away_qb_epa_ntile = qb_epa_ntile.y
    ,away_success_rt_ntile = success_rt_ntile.y
    
    ,r1_epa_pp_diff = r1_epa_pp.x - r1_epa_pp.y
    ,r2_epa_pp_diff = r2_epa_pp.x - r2_epa_pp.y
    ,r4_epa_pp_diff = r4_epa_pp.x - r4_epa_pp.y
    ,r4_pass_epa_diff = r4_pass_epa.x - r4_pass_epa.y
    ,r4_rush_epa_diff = r4_rush_epa.x - r4_rush_epa.y
    ,r4_qb_epa_diff = r4_qb_epa.x - r4_qb_epa.y
    ,r4_success_rt_diff = r4_success_rt.x - r4_success_rt.y
    ,r12_epa_pp_diff = r12_epa_pp.x - r12_epa_pp.y
    ,r12_pass_epa_diff = r12_pass_epa.x - r12_pass_epa.y
    ,r12_rush_epa_diff = r12_rush_epa.x - r12_rush_epa.y
    ,r12_qb_epa_diff = r12_qb_epa.x - r12_qb_epa.y
    ,r12_success_rt_diff = r12_success_rt.x - r12_success_rt.y
    ,r16_epa_pp_diff = r16_epa_pp.x - r16_epa_pp.y
    ,r16_pass_epa_diff = r16_pass_epa.x - r16_pass_epa.y
    ,r16_rush_epa_diff = r16_rush_epa.x - r16_rush_epa.y
    ,r16_qb_epa_diff = r16_qb_epa.x - r16_qb_epa.y
    ,r16_success_rt_diff = r16_success_rt.x - r16_success_rt.y
    ,epa_pp_ntile_diff = epa_pp_ntile.x - epa_pp_ntile.y
    ,pass_epa_ntile_diff  = pass_epa_ntile.x - pass_epa_ntile.y
    ,rush_epa_ntile_diff  = rush_epa_ntile.x - rush_epa_ntile.y
    ,qb_epa_ntile_diff  = qb_epa_ntile.x - qb_epa_ntile.y
    ,success_rt_ntile_diff  = success_rt_ntile.x - success_rt_ntile.y
    
    # MATCHUP METADATA
    ,weekday = weekday.x
    ,gameday = gameday.x
    ,gametime = gametime.x
    ,div_game = div_game.x
    ,home_rest = team_rest.x 
    ,away_rest = opponent_rest.x 
    ,rest_diff = rest_diff.x
  )
```

### Add Fivethirtyeight Elo scores and probabilities

```{r}
fullElo <- read_csv("https://projects.fivethirtyeight.com/nfl-api/nfl_elo.csv")
tail(fullElo)

fte <- fullElo %>%
  filter(season >= 2000) %>%
  dplyr::mutate(
    team1 = gsub("WSH", "WAS", team1),
    team2 = gsub("WSH", "WAS", team2),
    team1 = gsub("LAR", "LA", team1),
    team2 = gsub("LAR", "LA", team2),
    team1 = gsub('OAK','LV',team1),
    team2 = gsub('OAK','LV',team2),
    team1 = gsub('STL','LA',team1),
    team2 = gsub('STL','LA',team2),
    team1 = gsub('SD','LAC',team1),
    team2 = gsub('SD','LAC',team2),
    gameday = lubridate::as_date(date)
  ) %>%
  select(gameday,season,team1,team2,elo1_pre,elo2_pre,elo_prob1) %>%
  as_tibble()

gm_df <- game_w_features %>% 
  filter(week <= 17 | (season >= '2021' & week >= '18')) %>%
  dplyr::mutate(
    home_team = gsub('OAK','LV',home_team),
    away_team = gsub('OAK','LV',away_team),
    home_team = gsub('STL','LA',home_team),
    away_team = gsub('STL','LA',away_team),
    home_team = gsub('SD','LAC',home_team),
    away_team = gsub('SD','LAC',away_team),
    gameday = lubridate::as_date(gameday))

game_w_features <- gm_df %>%
  filter(week <= 17 | (season == '2021' & week >= '18')) %>%
  # mutate(gameday = lubridate::as_date(gameday)) %>%
  left_join(fte, by=c('gameday'='gameday','home_team'='team1','away_team'='team2',"season"))

# spit out a CSV 
write.csv(game_w_features,paste0("~/Documents/Scripts/R/pred/game_w_features.csv"), row.names = TRUE)
```
